Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-06-05 testbed.tex
fixed testbed intro

This section will present a testbed of thirty \ac{Raspi}s.
%Twenty \ac{Raspi} 1 model B revision 2 and ten \ac{Raspi} 2 model B V1.1.
They are all equipped with a memory card and are connected to the
university network by wire using their ethernet interface (eth0).

To easily discover the \ac{Raspi}s within the university network, they have all
been assigned a unique name under a common domain.
This eliminates the need to remember their individual \ac{IP}s.

A sketch of the testbed is depicted in figure~\ref{fig:testbed_setup}.
We have two types of \ac{Raspi}s in our setup.
Table \ref{tab:rasp_naming} shows the naming convention.

\begin{table}[ht!]
  \centering
  \begin{tabular}{|c|c|}
    \hline
      \textbf{Hostname} & \textbf{Model} \\ \hline
    Rasp00 - Rasp19 & Raspberry Pi 1 model B rev. 2 \\ \hline
    Rasp100 - Rasp109 & Rasperry Pi 2 model B V1.1 \\ \hline
  \end{tabular}
  \caption{Raspberry Pi naming convention}
  \label{tab:rasp_naming}
\end{table}

Accessing them is easily done using \ac{SSH}.
To address them within the university network,
The \ac{Raspi}s

The \ac{Raspi}s are all equipped with an ethernet interface (eth0) and
connected by wire to the university's \ac{LAN}. The university network
administration was provided
the \ac{MAC} addresses of the ethernet interfaces to automatically assign
the devices the desired \ac{IP} addresses as shown in the figure. In a personal
\ac{LAN}, another option could be to manually provide the devices with static
\ac{IP}s.

This section will be used to present a \ac{Raspi} testbed and the steps to
install and configure its devices.

The testbed is depicted in figure~\ref{fig:testbed_setup}. It consists of twenty
Raspberry Pi 1 model B rev 2 (Rasp01 to Rasp20) and
ten Raspberry Pi 2 model b V1.1 (Rasp2\_01 to Rasp2\_10).

The goal is to configure all the \ac{Raspi}s identically with the same software
and configurations. There are however a few differences between them that we
need to take into account. E.g. hostname and network addresses.

The testbed will be used by students that might need to change how the
devices are work either temporary or permanently. We will illustrate how this
can be accomplished using an overlaying filesystem.

Finally, it should be possible to store data that are generated from
simulations on the memory cards that are in the individual \ac{Raspi}.
%We using a overlaying filesystem that
%We will illustrate
%how that can be done. In the case of temporary changes, it is important that
%all changes can be easily reverted. We put a \ac{RAM} filesystem on top of
%the the root filesystem to .
%the configure the
%devices to behave according to their test specifications. We
%will also present how that might be done.

@@

The \ac{Raspi}s are all equipped with an ethernet interface (eth0) and
connected by wire to the university's \ac{LAN}. The university network
administration was provided
the \ac{MAC} addresses of the ethernet interfaces to automatically assign
the devices the desired \ac{IP} addresses as shown in the figure. In a personal
\ac{LAN}, another option could be to manually provide the devices with static
\ac{IP}s.
%that are
%assigned the \ac{IP} addresses in the figure and connected wired to the
%university's \ac{LAN}.

We use standard usb power adapters to power the \ac{Raspi}s.

@@

% Download and unpack image
\begin{lstlisting}[]
$ mkdir -p ${WORKDIR}
$ cd ${WORKDIR}
$ wget http://downloads.raspberrypi.org/raspbian_lite/images/raspbian_lite-2016-05-31/${IMAGE}.zip
$ unzip ${IMAGE}.zip
\end{lstlisting}
\FloatBarrier

>>>>

%This section will present a testbed of thirty \ac{Raspi}s and a \ac{PC}
The testbed consists of thirty \ac{Raspi}s that will be setup and controlled
by a testbed administrator from a \ac{PC}.
%Twenty \ac{Raspi} 1 model B revision 2 and ten \ac{Raspi} 2 model B V1.1.
The \ac{Raspi}s are each equipped with a memory card and connected to the
university network by wire using their ethernet interface (eth0).

To easily discover the \ac{Raspi}s within the university network,
they have all been assigned a unique name under a common domain.
This eliminates the need to remember their individual \ac{IP} addresses.

A sketch of the testbed is depicted in Figure~\ref{fig:testbed_setup}.
We have two types of \ac{Raspi}s in our setup.
Table~\ref{tab:rasp_naming} shows the naming convention.

\begin{table}[ht!]
  \centering
  \begin{tabular}{|c|c|}
    \hline
      \textbf{Hostname} & \textbf{Model} \\ \hline
    Rasp00 - Rasp19 & Raspberry Pi 1 model B rev. 2 \\ \hline
    Rasp20 - Rasp29 & Rasperry Pi 2 model B V1.1 \\ \hline
  \end{tabular}
  \caption{Raspberry Pi naming convention}
  \label{tab:rasp_naming}
\end{table}

%Accessing them is easily done using \ac{SSH}.
%To address them within the university network,
%The \ac{Raspi}s

%The \ac{Raspi}s are all equipped with an ethernet interface (eth0) and
%connected by wire to the university's \ac{LAN}. The university network
%administration was provided
%the \ac{MAC} addresses of the ethernet interfaces to automatically assign
%the devices the desired \ac{IP} addresses as shown in the figure. In a personal
%\ac{LAN}, another option could be to manually provide the devices with static
%\ac{IP}s.

%This section will be used to present a \ac{Raspi} testbed and the steps to
%install and configure its devices.

%The testbed is depicted in figure~\ref{fig:testbed_setup}. It consists of twenty
%Raspberry Pi 1 model B rev 2 (Rasp01 to Rasp20) and
%ten Raspberry Pi 2 model b V1.1 (Rasp2\_01 to Rasp2\_10).

The testbed should be usable by students to perform simulations and
measurements as part of their semester projects. It is therefore desired
that the \ac{Raspi}s are configured identically, but still allows the
students to alter the configurations.

When working on a testbed, it is not always desired to make persistant
changes to the devices. We also present how the concept of stacked
filesystems can be used to enable the means of both persistant and
temporary storage.

%It is therefore
%mandatory that the \ac{Raspi}s are easily configurable but also
%desired to

%The testbed should be usable by students that should be able to configure
%the \ac{Raspi}'s as desired. It is therefore

%The testbed will be used by students that might need to change how the
%devices are work either temporary or permanently. We will illustrate how this
%can be accomplished using an overlaying filesystem.

%The goal is to configure all the \ac{Raspi}s identically with the same software
%and configurations. There are however a few differences between them that we
%need to take into account. E.g. hostname and network addresses.

%Finally, it should be possible to store data that are generated from
%simulations on the memory cards that are in the individual \ac{Raspi}.
%We using a overlaying filesystem that
%We will illustrate
%how that can be done. In the case of temporary changes, it is important that
%all changes can be easily reverted. We put a \ac{RAM} filesystem on top of
%the the root filesystem to .
%the configure the
%devices to behave according to their test specifications. We
%will also present how that might be done.

@@

%The \ac{Raspi}s are all equipped with an ethernet interface (eth0) and
%connected by wire to the university's \ac{LAN}. The university network
%administration was provided
%the \ac{MAC} addresses of the ethernet interfaces to automatically assign
%the devices the desired \ac{IP} addresses as shown in the figure. In a personal
%\ac{LAN}, another option could be to manually provide the devices with static
%\ac{IP}s.
%that are
%assigned the \ac{IP} addresses in the figure and connected wired to the
%university's \ac{LAN}.

%We use standard usb power adapters to power the \ac{Raspi}s.

@@

% Download and unpack image
\begin{lstlisting}[]
$ mkdir -p ${WORKDIR}
$ cd ${WORKDIR}
$ wget http://kom.aau.dk/project/TuneSCode/raspi/${IMAGE}.zip
$ unzip ${IMAGE}.zip
\end{lstlisting}
\FloatBarrier

A more recent version of Raspbian lite may be available on 
\url{http://downloads.raspberrypi.org/raspbian_lite/images}.
